A template for the \emph{arxiv} style

### 2021-08-04 main/main.Rmd
draft on existing spatio + temporal R pkgs

@@

# Literature Review

Existing R pkgs for spatio/ temporal data

- `sp` + `spacetime` (time series data from package `xts`)

- `sf` 
  - simple features geometry 
  - only spatio
  - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure

- `stars`
  - a newer structure allow both spatio + temporal 
  - array

>>>>

<!-- While satellite imagery can record information on a continuous spatial domain, other spatial information are recorded at a fixed set of locations characterised by latitude, longitude, and other spatial metadata. Examples of this type of data includes climate measures from weather observation stations and river level data recorded by electronic gauges. Analysing this data requires less considerations on the geometry type and map projection while more on how measures in these fixed locations changes across the time domain and whether these changes are related for adjacent locations. -->

@@

# Existing data structure for spatio and temporal data

There has been a large class of implementations dedicated to processing the spatial data. This includes `sf` [@pebesma2018simple] and its precedent `sp` [@pebesma2005s] for ... and `raster` [@raster] and `terra` [@terra] for raster data. While these implementations specialised in geographic manipulations with different type of simple features, it doesn't incorporate a temporal dimension in the data structure. Project like `stars` [@stars] and `spacetime` [@spacetimebook] by R-Spatial allows for both space and time dimension for raster and vector data, but the underlying data structure is a multi-dimensional array, which could be difficult to operate for R users who are more familiar with the operation in 2D dataframe/ tibble.

<!-- - `sp` + `spacetime` (time series data from package `xts`) -->

<!-- - `sf`  -->
<!--   - simple features geometry  -->
<!--   - only spatio -->
<!--   - good for geographical but these geometry types i.e. multipolygon, can be hard to manipulate. Would be nice to have a vector data structure -->

<!-- - `stars` -->
<!--   - a newer structure allow both spatio + temporal  -->
<!--   - array -->

In the temporal aspect, the `tsibble` [@tsibbles] structure and its tidyverts ecosystem have provided a [... ] workflow to work with temporal data. In a tsibble structure, temporal data is characterised by `index` and `key` where `index` is the temporal identifier and `key` is the identifier for multiple series, which could be used as a spatio identifier. However, a tsibble object, by construction, always requires the `index` in its structure. This makes it less appealing for spatio-temporal data since the output of calculated spatio-specific variables (i.e. features of each series) don't have the time dimension. Analysts will either need to have an additional step to join this output to the original tsibble or operate with variables stored in two separate objects. In addition, the long form structure of a tsibble object means spatio variables (i.e. longitude, latitude, and features of each series if joined back to the tsibble) of each spatio identifier will be repetitively recorded at each timestamp. This repetition is unnecessary and would inflate the object size for long series.

### 2021-08-04 main/main.tex
draft on existing spatio + temporal R pkgs

% Pandoc citation processing

@@

\hypertarget{literature-review}{%
\section{Literature Review}\label{literature-review}}

Existing R pkgs for spatio/ temporal data

\begin{itemize}
\item
  \texttt{sp} + \texttt{spacetime} (time series data from package
  \texttt{xts})
\item
  \texttt{sf}

  \begin{itemize}
  \tightlist
  \item
    simple features geometry
  \item
    only spatio
  \item
    good for geographical but these geometry types i.e.~multipolygon,
    can be hard to manipulate. Would be nice to have a vector data
    structure
  \end{itemize}
\item
  \texttt{stars}

  \begin{itemize}
  \tightlist
  \item
    a newer structure allow both spatio + temporal
  \item
    array
  \end{itemize}
\end{itemize}

@@

\hypertarget{summary}{%
\section{Summary}\label{summary}}

>>>>

% Pandoc citation processing
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1 \everypar{\setlength{\hangindent}{\cslhangindent}}\ignorespaces\fi
  % set entry spacing
  \ifnum #2 > 0
  \setlength{\parskip}{#2\baselineskip}
  \fi
 }%
 {}
\usepackage{calc} % for calculating minipage widths
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\hypertarget{existing-data-structure-for-spatio-and-temporal-data}{%
\section{Existing data structure for spatio and temporal
data}\label{existing-data-structure-for-spatio-and-temporal-data}}

There has been a large class of implementations dedicated to processing
the spatial data. This includes \texttt{sf} (E. J. Pebesma 2018) and its
precedent \texttt{sp} (E. Pebesma and Bivand 2005) for \ldots{} and
\texttt{raster} (Hijmans 2020) and \texttt{terra} (Hijmans 2021) for
raster data. While these implementations specialised in geographic
manipulations with different type of simple features, it doesn't
incorporate a temporal dimension in the data structure. Project like
\texttt{stars} (E. Pebesma 2021) and \texttt{spacetime} (Bivand,
Pebesma, and Gomez-Rubio 2013) by R-Spatial allows for both space and
time dimension for raster and vector data, but the underlying data
structure is a multi-dimensional array, which could be difficult to
operate for R users who are more familiar with the operation in 2D
dataframe/ tibble.

In the temporal aspect, the \texttt{tsibble} (Wang, Cook, and Hyndman
2020) structure and its tidyverts ecosystem have provided a {[}\ldots{}
{]} workflow to work with temporal data. In a tsibble structure,
temporal data is characterised by \texttt{index} and \texttt{key} where
\texttt{index} is the temporal identifier and \texttt{key} is the
identifier for multiple series, which could be used as a spatio
identifier. However, a tsibble object, by construction, always requires
the \texttt{index} in its structure. This makes it less appealing for
spatio-temporal data since the output of calculated spatio-specific
variables (i.e.~features of each series) don't have the time dimension.
Analysts will either need to have an additional step to join this output
to the original tsibble or operate with variables stored in two separate
objects. In addition, the long form structure of a tsibble object means
spatio variables (i.e.~longitude, latitude, and features of each series
if joined back to the tsibble) of each spatio identifier will be
repetitively recorded at each timestamp. This repetition is unnecessary
and would inflate the object size for long series.

@@

\hypertarget{summary}{%
\section*{Summary}\label{summary}}
\addcontentsline{toc}{section}{Summary}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-spacetimebook}{}%
Bivand, Roger S., Edzer Pebesma, and Virgilio Gomez-Rubio. 2013.
\emph{Applied Spatial Data Analysis with {R}, Second Edition}. Springer,
NY. \url{https://asdar-book.org/}.

\leavevmode\hypertarget{ref-raster}{}%
Hijmans, Robert J. 2020. \emph{Raster: Geographic Data Analysis and
Modeling}. \url{https://CRAN.R-project.org/package=raster}.

\leavevmode\hypertarget{ref-terra}{}%
---------. 2021. \emph{Terra: Spatial Data Analysis}.
\url{https://CRAN.R-project.org/package=terra}.

\leavevmode\hypertarget{ref-stars}{}%
Pebesma, Edzer. 2021. \emph{Stars: Spatiotemporal Arrays, Raster and
Vector Data Cubes}. \url{https://CRAN.R-project.org/package=stars}.

\leavevmode\hypertarget{ref-pebesma2018simple}{}%
Pebesma, Edzer J. 2018. {``Simple Features for r: Standardized Support
for Spatial Vector Data.''} \emph{R J.} 10 (1): 439.

\leavevmode\hypertarget{ref-pebesma2005s}{}%
Pebesma, Edzer, and Roger S Bivand. 2005. {``S Classes and Methods for
Spatial Data: The Sp Package.''} \emph{R News} 5 (2): 9--13.

\leavevmode\hypertarget{ref-tsibbles}{}%
Wang, Earo, Dianne Cook, and Rob J Hyndman. 2020. {``A New Tidy Data
Structure to Support Exploration and Modeling of Temporal Data.''}
\emph{Journal of Computational and Graphical Statistics} 29 (3):
466--78. \url{https://doi.org/10.1080/10618600.2019.1695624}.

\end{CSLReferences}
